---
title: "Bcities"
author: "Issy Belzil & Mikaela Murgoitio"
date: '2017-10-24'
output: 
  html_document:
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r eval=FALSE}
#setwd("~/Documents/2DataViz /BCity")

library(ggplot2)
library(dplyr)
library(tidyr)
library(gudatavizfa17) #Load everytime

#install.packages('googleVis')
#install.packages("ggmap")
library(ggmap)
library(maps)
library(googleVis)

```



```{r setOptions, message=FALSE,echo=FALSE,warning=FALSE}
suppressPackageStartupMessages(library(knitr)) 
suppressPackageStartupMessages(library(googleVis)) 
library(ggplot2)
library(dplyr)
library(tidyr)
library(gudatavizfa17) #Load everytime

#install.packages('googleVis')
#install.packages("ggmap")
library(ggmap)
library(maps)

opts_chunk$set(cache=TRUE, message=FALSE,size = 'footnotesize',warning=FALSE,results='asis')
op <- options(gvis.plot.tag='chart') 
```

###Question 1: Please look at relationships between different variables and identify variables that appear to be strongly related to having a higher rank in the survey.

Facet Graphs: First, we looked at all the variables and compared them against the rankings to try to find relationships.


```{r fig1, fig.height = 6, fig.width = 12}
facetcities=gather(bcities, "Variable", "Values", 1,3:13)

facetcities=facetcities[facetcities$Variable!="City",]


facetcities$Values=as.numeric(facetcities$Values)

ggplot(facetcities,aes(x=Rank,y=Values)) + geom_point(size=.5)+ facet_wrap(~Variable, scales = "free")+geom_smooth(method="lm") + labs(list(title = "Variables and their Influence on the City Ranking", x = "Rank", y = "Variable"))
```

This facet graph gives us a an overview at the relationships between the ranking of the cities and the variables:  Population, Bars, Restaurants, Museums, Libraries, Number of Pro Sports Teams, Park Acres to 1000 residents, Colleges, Percentage of people with a Graduate Degree, Median Household Incomes and Percent Unemployed. From this, we wanted to examine some of the relationships further. Then, we ran correlation tests with the variables that we saw the highest relationship with, and the variables we were unsure about. 

#####Comparing Rank with Percentage of People with Graduate Degrees

```{r}
ggplot(bcities,aes(x=Rank,y=Percent.with.Graduate.Degree)) + geom_point(shape=1)  +
geom_smooth(method="lm")

#converting Rank to numeric value

bcities$Rank <- as.numeric(bcities$Rank)
str(bcities)

#do the correlation command 

cor(bcities$Rank, y = bcities$Percent.with.Graduate.Degree, use = "everything",
    method = c("pearson"))

```

After running a pearson correlation test, the result of -0.53 proves that there is a moderate negative relationship between Rank and Percent with Graduate Degree.  

#####Comparing Rank with Percentage of People who are Unemployed

```{r}

ggplot(bcities,aes(x=Rank,y=Percent.unemployed)) + geom_point(shape=1)  +
geom_smooth(method="lm")

#Performing Correlation Command

cor(bcities$Rank, y = bcities$Percent.unemployed, use = "everything",
    method = c("pearson"))

```

To prove there is not a relationship between unemployment and the ranking of the cities, we ran another correlation test between the two variables. The r coefficient = 0.0293, proving no relationship.


#####Comparing Rank with Number of Colleges 

```{r}
ggplot(bcities,aes(x=Rank,y=Colleges)) + geom_point(shape=1)  +
geom_smooth(method="lm")

#converting Colleges to numeric value

bcities$Colleges <- as.numeric(bcities$Colleges)

#Performing Correlation Command

cor(bcities$Rank, y = bcities$Colleges, use = "everything",
    method = c("pearson"))
```

Rank and Colleges correlation of = -0.272. Thus, there is a weak downhill correlation between cities ranks and how many colleges they have. 




#####Comparing Rank with Number of Pro Sports Team 

```{r}
ggplot(bcities,aes(x=Rank,y=Pro.Sports.Teams)) + geom_point(shape=1)  +
geom_smooth(method="lm")

#converting Pro.Sports.Teams to numeric value

bcities$Colleges <- as.numeric(bcities$Pro.Sports.Teams)

#Performing Correlation Command

cor(bcities$Rank, y = bcities$Pro.Sports.Teams, use = "everything",
    method = c("pearson"))
```

Rank and Colleges correlation of = -0.222. Thus, there is a weak downhill correlation between how rank and how many pro sports teams. 




#####Comparing Rank with Number of Museums 

```{r}
ggplot(bcities,aes(x=Rank,y=Museums)) + geom_point(shape=1)  +
geom_smooth(method="lm")

#converting Museums to numeric value

bcities$Museums <- as.numeric(bcities$Museums)

#Performing Correlation Command

cor(bcities$Rank, y = bcities$Museums, use = "everything",
    method = c("pearson"))
```

Observing the correlation between Ranking of the City and the amount of museums in a city, it is a larger correlation than Colleges and Ranking. Although the correlation is still weak at r = -0.288.

#####Comparing Rank with Number of Libraries

```{r}
ggplot(bcities,aes(x=Rank,y=Libraries)) + geom_point(shape=1)  +
geom_smooth(method="lm")

#converting Libraries to numeric value

bcities$Libraries <- as.numeric(bcities$Libraries)

#Performing Correlation Command

cor(bcities$Rank, y = bcities$Libraries, use = "everything",
    method = c("pearson"))
```

Weak correlation at r = -0.247

#####Comparing Rank with Number of Bars 

```{r}
ggplot(bcities,aes(x=Rank,y=Bars)) + geom_point(shape=1)  +
geom_smooth(method="lm")

#converting Bars to numeric value

bcities$Bars <- as.numeric(bcities$Bars)

#Performing Correlation Command

cor(bcities$Rank, y = bcities$Bars, use = "everything",
    method = c("pearson"))
```

Low correlation. Weak downhill correlation. 

#####Comparing Rank with Population

```{r}
ggplot(bcities,aes(x=Rank,y=Population)) + geom_point(shape=1)  +
geom_smooth(method="lm")

#converting Population to numeric value

bcities$Population <- as.numeric(bcities$Population)

#Performing Correlation Command

cor(bcities$Rank, y = bcities$Population, use = "everything",
    method = c("pearson"))
```

One of the lowest correlations, found thus far between Ranking of City and a factor that may have thought to be influenced by the ranking. The pearson correlation variable = -0.0929. This is surprising as the top 10 have higher than average populations, but overall, there is no correlation. Having a higher population will not effect your ranking. 

#####Comparing Rank with Number of Restaurants 

```{r}
ggplot(bcities,aes(x=Rank,y=Restaurants)) + geom_point(shape=1)  +
geom_smooth(method="lm")

#converting Restaurants to numeric value

bcities$Restaurants <- as.numeric(bcities$Restaurants)

#Performing Correlation Command

cor(bcities$Rank, y = bcities$Restaurants, use = "everything",
    method = c("pearson"))
```

There is weak to no relationship between having a large number of restaurants and having a good ranking, with the pearson correlation at -0.167. 


#####The factor with the highest correlation to Rank is people with Graduate Degrees and has the strongest relationship. 


### Question 2: Plot the cities on a map of the US.
```{r}
require(maps)
require(ggmap)
```

Creating columns to better display data(cityrank) and latlong variable to combine coordinates for plotting on gvisgeochart:
```{r}
bcities$latlong=paste(bcities$lat,bcities$lon, sep=":")
bcities$cityrank=paste(bcities$City,bcities$Rank,sep=": Rank ")
```
Plot bcities data using gvisGeoChart:
```{r}
CityLoc=gvisGeoChart(bcities,"latlong",hovervar = "cityrank",options=list(displayMode="Markers",region="US"))
plot(CityLoc)
```


###Question 3: Identify the States in which these cities lie and then develop a choropleth that colors different states based on the number of cities that are present in each State.

Manually input State names:

```{r}
bcities$States = c("California", "Alaska", "Nebraska", "Missouri", "Ohio", "Virginia", "Arizona", "Arizona", "Nevada", "Texas", "Virginia", "North Carolina", "Oklahoma", "Indiana", "Colorado", "Florida", "Nebraska", "California", "New York", "California", "Texas", "Maryland", "Illinois", "Washington", "Wisconsin", "Kentucky", "Pennsylvania", "Oklahoma", "Texas", "Ohio", "Ohio", "Hawaii", "North Carolina", "Wisconsin", "Georgia", "Kansas", "Lousiana", "Tennessee", "Minnesota", "Pennsylvania", "Minnesota", "California", "Texas", "New York", "Colorado", "Oregon", "Massachusetts", "District of Columbia", "Washington", "California")
```
Create Data set of States and count cities within each state. Convert char variable to factor variable, and int to number variable. 
```{r}
bstatescount = bcities %>% group_by(States) %>% summarise(Count = n())
bstatescount$Ccount <- as.numeric(bstatescount$Count)
bstatescount$States <- as.factor(bstatescount$States)
str(bstatescount)
```


Plot dataset using GvizGeoChart
```{r}

CityCount = gvisGeoChart(bstatescount,
                         locationvar = "States", 
                         colorvar = "Count", 
                         options=list(region="US", displayMode="regions",
                                      resolution="provinces", width=800,
                                      height=600))

plot(CityCount)

```


